### 2021-01-05 cover letter/resubmission.tex

\documentclass[a4paper]{article}
\input{../preamble/packages}
\usepackage{pdfpages}

\date{}

\title{Resubmission \\ \vspace{1cm} \textbf{\Large Simultaneous Elicitation of Scoring Rule and Agent Preferences for Robust Winner Determination} \\ {\large Submission ID: 554, AAMAS-2021}}

\begin{document}
\maketitle

\section{Cover Letter}
\input{cover_letter.tex}

\section{Reviews AAMAS-2021}
\input{reviews_aamas.tex}

\section{Reviews IJCAI-20}
\input{reviews_ijcai.tex}

\section{Rejected paper AAMAS-2021}
\includepdf[pages=-]{aamas21_paper_554.pdf}

\end{document}