### 2021-01-09 cover letter/resubmission.tex
Canvas

\RequirePackage[l2tabu, orthodox]{nag}
\documentclass[version=3.21, pagesize, twoside=off, bibliography=totoc, DIV=calc, fontsize=12pt, a4paper]{scrartcl}
\input{preamble/packages}
\input{preamble/redac}
\input{preamble/math_basics}
\input{preamble/math_mine}
\usepackage{pdfpages}

%I find these settings useful in draft mode. Should be removed for final versions.
	%Which line breaks are chosen: accept worse lines, therefore reducing risk of overfull lines. Default = 200.
		\tolerance=2000
	%Accept overfull hbox up to...
		\hfuzz=2cm
	%Reduces verbosity about the bad line breaks.
		\hbadness 5000
	%Reduces verbosity about the underful vboxes.
		\vbadness=1300

\title{Resubmission: Simultaneous Elicitation of Scoring Rule and Agent Preferences for Robust Winner Determination}
\author{Submission ID: 554, AAMAS-2021}
\date{}
\hypersetup{
	pdfsubject={Cover letter},
	pdfkeywords={elicitation},
}

\begin{document}
\maketitle

\section{Cover Letter}
\input{cover_letter.tex}

\section{Reviews AAMAS-2021}
\input{reviews_aamas.tex}

\section{Reviews IJCAI-20}
\input{reviews_ijcai.tex}

\section{Rejected paper AAMAS-2021}
%\includepdf[pages=-]{aamas21_paper_554.pdf}

%\bibliography{bibl}

\end{document}